---
title: "PS8"
author: "Prateek Kumar"
date: "3 November 2018"
output: word_document
---

#Problem 1
```{r}
training <- rep(c(1,1,1,0,0,0,-1,-1,-1),each=3)
training
feedback <- rep(c(1,0,-1,1,0,-1,1,0,-1),each=3)
feedback
```
We have two 3-level independent variables: 
* training (no training, standard training, interactive training) 
* feedback (no feedback, feedback on errors, feedback on correct)

Correlation between training and feedback
```{r}
cor(training, feedback)
```
Dot product between training and feedback
```{r}
training %*% feedback
```
We can verify that both of them are uncorrelated and orthogonal since their correlation and dot product is zero.

#Create one that is orthogonal to the two existing IVs, and verify that it is orthogonal
```{r}
group <- rep(c(1,-1,0,-1,0,1,0,1,-1),each=3)
group
```
Here the above IV "group" is both uncorrelated and orthogonal to both training and feedback, which we verified below

Correlation between training and group
```{r}
cor(training, group)
```
Dot product between training and group
```{r}
training %*% group
```
Its orthogonal and uncorrelated with training

Correlation between feedback and group
```{r}
cor(feedback, group)
```
Dot product between feedback and group
```{r}
feedback %*% group
```
Its orthogonal and uncorrelated with feedback as well.


#Which of the following grouping variables are both orthogonal and uncorrelated with both feedback and train
```{r}
set.seed(100)
group1 <- 1:27
group2 <- rep(-1:1, 9)
group3 <- 1:27-mean(1:27)
group4 <- rep(1:3, 9)
group5 <- rep(-1:1,each=9)
group6 <- runif(27)-.5
group7 <- rep(1:9,each=3)
```
We will calculate the correlation and dot product of each group IV with training and feedback

Firstly calculating with training:

Correlation between training and group1
```{r}
cor(training, group1)
```
Dot product between training and group1
```{r}
training %*% group1
```

Correlation between training and group2
```{r}
cor(training, group2)
```
Dot product between training and group2
```{r}
training %*% group2
```

Correlation between training and group3
```{r}
cor(training, group3)
```
Dot product between training and group3
```{r}
training %*% group3
```

Correlation between training and group4
```{r}
cor(training, group4)
```
Dot product between training and group4
```{r}
training %*% group4
```

Correlation between training and group5
```{r}
cor(training, group5)
```
Dot product between training and group5
```{r}
training %*% group5
```

Correlation between training and group6
```{r}
cor(training, group6)
```
Dot product between training and group6
```{r}
training %*% group6
```

Correlation between training and group7
```{r}
cor(training, group7)
```
Dot product between training and group7
```{r}
training %*% group7
```

We see that only group2 and group4 are uncorrelated and orthogonal with training because the correlation and dot product is zero in this case

Now calculating with feedback

Correlation between feedback and group1
```{r}
cor(feedback, group1)
```
Dot product between feedback and group1
```{r}
feedback %*% group1
```

Correlation between feedback and group2
```{r}
cor(feedback, group2)
```
Dot product between feedback and group2
```{r}
feedback %*% group2
```

Correlation between feedback and group3
```{r}
cor(feedback, group3)
```
Dot product between feedback and group3
```{r}
feedback %*% group3
```

Correlation between feedback and group4
```{r}
cor(feedback, group4)
```
Dot product between feedback and group4
```{r}
feedback %*% group4
```

Correlation between feedback and group5
```{r}
cor(feedback, group5)
```
Dot product between feedback and group5
```{r}
feedback %*% group5
```

Correlation between feedback and group6
```{r}
cor(feedback, group6)
```
Dot product between feedback and group6
```{r}
feedback %*% group6
```

Correlation between feedback and group7
```{r}
cor(feedback, group7)
```
Dot product between feedback and group7
```{r}
feedback %*% group7
```

We see that group2, group4 and group5 are uncorrelated and orthogonal with feedback because the correlation and dot product is zero in this case

So we conclude from the above results that both group2 and group4 are uncorrelated and orthogonal with both training and feedback.

#Problem2
```{r}
data <- data.frame(buyerid = group7,
                   timeframe=group4)

set.seed(103)

buyerbase <- runif(9)
timebase <- c(1,2,2.5)
agebase  <- 20+runif(9)* 40

data$age <- round(agebase[data$buyerid])
data$purchase <- round( 25 + buyerbase[data$buyerid] * 50 +
                          data$age * .5+ 
                          timebase[data$timeframe] * 7 +
                          rnorm(27)*4,2)


lm(purchase ~ timeframe+age, data = data)$coef
lm(purchase ~ timeframe, data = data)$coef
lm(purchase ~ age, data = data)$coef

age1 <- data$age-mean(data$age)
tf1 <- data$timeframe-mean(data$timeframe)

lm(purchase ~ tf1+age1, data = data)$coef
lm(purchase ~ tf1, data = data)$coef
lm(purchase ~ age1, data = data)$coef

data2 <- data[-1,]
lm(purchase ~ timeframe+age, data = data2)$coef
lm(purchase ~ timeframe, data = data2)$coef
lm(purchase ~ age, data = data2)$coef

age2 <- data2$age-mean(data2$age)
tf2 <- data2$timeframe-mean(data2$timeframe)

lm(purchase ~ tf2+age2, data = data2)$coef
lm(purchase ~ tf2, data = data2)$coef
lm(purchase ~ age2, data = data2)$coef
```

#Problem 3
#Q1
```{r}
down <- read.csv("DownloadData.csv")
down$Month <- as.factor(substr(down$Date,6,8))
down$MonthNumber <- 1:nrow(down)

plot(as.numeric(down$Date),down$Downloads,xaxt="n",bty="n",pch=21,cex=.5,type="p",las=1,
     ylab="Downloads",xlab="Year")
axis(1,0:12*12,2006:2018,las=3,cex.axis=.95)

lines(lowess(down$Downloads~down$MonthNumber, f=0.10), col="red")
lines(lowess(down$Downloads~down$MonthNumber, f=0.25), col="green")
lines(lowess(down$Downloads~down$MonthNumber, f=0.50), col="blue")
lines(lowess(down$Downloads~down$MonthNumber, f=0.01), col="purple")
lines(lowess(down$Downloads~down$MonthNumber, f=0.05), col="gold")


plot(as.numeric(down$Date),down$Downloads,xaxt="n",bty="n",pch=21,cex=.5,type="p",las=1,
     ylab="Downloads",xlab="Year")
axis(1,0:12*12,2006:2018,las=3,cex.axis=.95)

lmodel <- loess(down$Downloads~down$MonthNumber)
xs <- 0:50000/100
points(xs,predict(lmodel,xs),type="l",col="red",lwd=2)
summary(lmodel)
print(paste("R^2:", cor(down$Downloads,predict(lmodel))^2))

lmodel2 <- loess(down$Downloads~down$MonthNumber,span=.2,degree = 2)
points(xs,predict(lmodel2,xs),type="l",col="green",lwd=2)
summary(lmodel2)
print(paste("R^2:", cor(down$Downloads,predict(lmodel2))^2))

lmodel3 <- loess(down$Downloads~down$MonthNumber,span=.3)
points(xs,predict(lmodel3,xs),type="l",col="navy",lwd=2)
summary(lmodel3)
print(paste("R^2:", cor(down$Downloads,predict(lmodel3))^2))

lmodel4 <- loess(down$Downloads~down$MonthNumber,span=.4)
points(xs,predict(lmodel4,xs),type="l",col="gold",lwd=2)
summary(lmodel4)
print(paste("R^2:", cor(down$Downloads,predict(lmodel4))^2))
```

#Q2
```{r}
plot(down$MonthNumber,down$Downloads,xaxt="n",bty="n",pch=21,cex=.5,type="p",las=1,
     ylab="Downloads",xlab="Year")
axis(1,0:12*12,2006:2018,las=3,cex.axis=.95)


lines(lowess(down$Downloads~down$MonthNumber))
poly2 <- lm(Downloads ~ poly(MonthNumber, 2), data = down)      
summary(poly2)

pred<-predict(poly2)


plot(down$MonthNumber,down$Downloads)
lines(down$MonthNumber,y=pred,col="red")

poly1 <- lm(Downloads ~ poly(MonthNumber, 1), data = down)
poly3 <- lm(Downloads ~ poly(MonthNumber, 3), data = down)
poly4 <- lm(Downloads ~ poly(MonthNumber, 4), data = down)
poly5 <- lm(Downloads ~ poly(MonthNumber, 5), data = down)


data.frame(model = paste ("lm" ,1:5 , sep =""),
               rbind ( extractAIC ( poly1 ),
                       extractAIC ( poly2 ),
                       extractAIC ( poly3 ),
                       extractAIC ( poly4 ),
                       extractAIC ( poly5 )))


extractBIC <- function (model)
{
  extractAIC (model ,k= log ( length ( model $ residuals )))
}

data.frame( model = paste ("lm" ,1:5 , sep =""),
            rbind ( extractBIC ( poly1 ),
                    extractBIC ( poly2 ),
                    extractBIC ( poly3 ),
                    extractBIC ( poly4 ),
                    extractBIC ( poly5 )))
```

#Q3
```{r}
fit3 <- lm(Downloads ~ poly(MonthNumber,2)+as.factor(Month)+0, data = down)  
pred<-predict(fit3)
plot(down$MonthNumber,down$Downloads)
lines(down$MonthNumber,y=pred,col="red")

```