---
title: "PS6"
author: "Prateek Kumar"
date: "20 October 2018"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r}
set.seed(1000)
indoor<- sample(c("A","B","C","D"),prob=c(5,8,25,6),size=200,replace=T)
outdoor <-sample(c("A","B","C","D"),prob=c(5,20,3,6),size=200,replace=T)
tab1 <- table(indoor,outdoor)                              #cross-tabulations of each pair of responses
tab2 <- table(c(indoor,outdoor),rep(c("I","O"),each=200))  #what responses were given for indoor and outdoor across the four options

row.names(tab1) <- c('clowns','horses','chocolate fountain','magician') #Changing the row and column names
colnames(tab1) <- c('clowns','horses','chocolate fountain','magician')
row.names(tab2) <- c('clowns','horses','chocolate fountain','magician')
addmargins(tab1)
tab2
```

Here we have two tables: tab1 and tab2, tab1 compares the children indoor response across various outdoor responses so its a cross tabulation of each pair of responses and tab2 gives us the details as what responses were given for both indoor and outdoor across the four options.

#Determine whether responses to indoor preference influenced individuals response to outdoor preference
```{r}
#chisq.test(tab2[,1],p = tab2[,2],rescale.p = T)
#install.packages("corrplot")
library(corrplot)
library(BayesFactor)

ct_t1 <- chisq.test(tab1) 
ct_t1 #Chi-squared test on tab1
bf_t1 <- contingencyTableBF(tab1, sampleType = 'indepMulti', fixedMargin = 'cols') 
bf_t1 #bayes factor contingency table tests for tab1
bf_t1@bayesFactor$bf  #bayes factor value
corrplot(ct_t1$residuals, is.cor = F)
```


#Determine whether indoor and outdoor preferences were the same
```{r}
ct_t2 <- chisq.test(tab2)
ct_t2 #Chi-squared test on tab2
bf_t2 <- contingencyTableBF(tab2, sampleType = 'indepMulti', fixedMargin = 'cols')
bf_t2 #bayes factor contingency table tests for tab2
bf_t2@bayesFactor$bf  #bayes factor value
corrplot(ct_t2$residuals, is.cor = F)
#Indoor high association for chocolate fountain(strong positive association) & high association for outdoor and horses
#There is repulsion between outdoor and chocolate fountain
```

#Q2

```{r}
cardat <- read.table(text="age gender type origin origin.last carval carval.last
                     34 F SUV US US 16400 15800
                     31 M Truck US Europe 16900 16000
                     47 M Sedan US US 18800 17100
                     21 F Sedan Japan Japan 16000 15500
                     42 M SUV US Japan 16800 16100
                     43 F SUV US US 17200 16300
                     60 F Truck Europe Europe 19900 17800
                     37 M Truck Europe Europe 17100 16200
                     46 F SUV Japan Japan 16900 16300
                     27 M Sedan US US 16200 15700
                     50 M SUV US US 18800 17100
                     64 F SUV Japan US 50700 31700
                     33 M SUV Japan Japan 16500 15900
                     39 M Truck US Europe 17000 16200
                     58 F Sedan Japan US 19400 17500
                     53 F SUV US Europe 19200 17400
                     29 F Sedan US Japan 16300 15700
                     37 F Sedan US US 17300 16300
                     37 M SUV US Japan 18200 16700
                     54 F Sedan Japan Japan 24500 19800
                     46 F SUV Japan Europe 18000 16700
                     55 F SUV US Japan 28900 21700
                     46 F Truck US Europe 16600 16100
                     57 M SUV Europe Europe 24300 19700
                     40 M SUV US US 16800 16100
                     27 M Sedan Japan US 16900 16000
                     58 M SUV Europe Europe 20300 17900
                     64 M Truck US US 40600 27100
                     47 M Truck US Europe 18400 16900
                     32 M Truck US US 15900 15600
                     43 F Sedan Japan US 17200 16300
                     66 M Truck Europe Europe 19100 17500
                     36 F SUV US Japan 16900 16100
                     68 M Truck US US 69300 40100
                     54 F Sedan Japan US 17000 16400
                     64 M Truck Japan Europe 34900 24600
                     27 M SUV Japan Europe 15800 15500
                     51 F Sedan Japan Japan 29000 21700
                     69 M Sedan US Japan 54400 33400
                     25 F Sedan Japan Japan 15800 15500",header=T)
```

#Is there an impact of gender on the type of car purchased?
```{r}
library(corrplot)
library(BayesFactor)

t1 <- table(cardat$gender, cardat$type)
t1 #extracting the table of gender w.r.t. car type

ct_1 <- chisq.test(t1) 
ct_1 #Chi-squared test on t1
bf_1 <- contingencyTableBF(t1, sampleType = 'indepMulti', fixedMargin = 'cols') 
bf_1 #bayes factor contingency table tests for t1
bf_1@bayesFactor$bf  #bayes factor value
corrplot(ct_1$residuals, is.cor = F)

```

#Is there a difference in amount paid for a car for men versus women?
```{r}
library(vioplot)
library(BayesFactor)

t2 <- table(cardat$carval,cardat$gender) 

women <- subset(cardat, cardat$gender == "F")
men <- subset(cardat, cardat$gender == "M")
#overall differences
t.test(women$carval, men$carval, alternative = 'less') #comparing two independent samples
```

```{r}
t2_bf <- ttestBF(women$carval, men$carval) #bayesfactor test
t2_bf
t2_bf@bayesFactor$bf
```

```{r}
vioplot(women$carval, men$carval, names = c('female','male'),col = c('gold'))
title('Violin plot for car values of men and women')
grid()
```

```{r}
mean(men$carval) #mean amout of car of men and women
mean(women$carval)

length(men$carval) #number of men and women
length(women$carval)
```

#Do people tend to buy vehicles from of the same origin as their last vehicle (US, europe, japan)? 
```{r}
new <- cardat$origin
old <- cardat$origin.last
t3 <- table(old,new)
addmargins(t3)

ct_3 <- chisq.test(t3) 
ct_3 #Chi-squared test on t3
bf_3 <- contingencyTableBF(t3, sampleType = 'indepMulti', fixedMargin = 'cols') 
bf_3 #bayes factor contingency table tests for t3
bf_3@bayesFactor$bf  #bayes factor value
corrplot(ct_3$residuals, is.cor = F)

```

#Is there a relationship between driver age and the value of his car?
```{r}
t4 <- cardat[,c('age','carval')] #extracting age and car value

model1 <- lm(t4$carval~t4$age) #automatic fit
model1

par(mfrow=c(1,1)) #plotting carval and age 
plot(t4$age,t4$carval,pch=16,cex=1.5,col="gold",
     main=paste("Best-fitting line\n", "y = ",round(model1$coef[1] ,2) ," + ", 
                round(model1$coef[2],3) , " * x\nPlot for age and carval",sep=""),
     type = 'p', xlab = 'Age', ylab = 'Car Value')

points(t4$age,t4$carval,pch=1,cex=1.5,col="grey20")
abline(model1$coef,lwd=2) #best fit line

summary(model1)

predictedy <- model1$coef[1] + model1$coef[2]* t4$age #predictions of value we did not see

plot(predictedy, t4$carval,cex=1,col="red",pch=16,
     main = 'Plot for predictedy and carval',xlab = 'predictedy', ylab = 'carval') #plotting them
points(predictedy,t4$carval,pch=1,cex=1,col="black")
abline(0,1)
cor(t4$carval,predictedy)^2
```

```{r}
#Bayes factor test
x4 <- ttestBF(t4$age,t4$carval,paired = T)
x4@bayesFactor$bf
```

#What is your best estimate for the value of a car driven by a 32, 52, and 62-year-old?
```{r}
length(cardat$carval[cardat$age==32])
length(cardat$carval[cardat$age==52])
length(cardat$carval[cardat$age==62])
```

#Is there a relationship between how much someone paid for their previous car and how much they paid for their current car?
```{r}
t6 <- cardat[,c("carval","carval.last")] #extracting age and car value
cor.test(t6$carval, t6$carval.last)
```
The correlation is almost perfect

```{r}
model2 <- lm(t6$carval~t6$carval.last) #automatic fit
model2

par(mfrow=c(1,1)) #plotting carval and age 
plot(t6$carval.last,t6$carval,pch=16,cex=1.5,col="gold",
     main=paste("Best-fitting line\n", "y = ",round(model1$coef[1] ,2) ," + ", 
                round(model1$coef[2],3) , " * x\nPlot for previous car value and current car value",sep=""),
     xlab = 'Previous car value', ylab = 'Current car value')

points(t6$carval.last,t6$carval,pch=1,cex=1.5,col="black")
abline(model2$coef,lwd=2) #best fit line

summary(model2)

predictedy <- model2$coef[1] + model2$coef[2]* t6$carval.last #predictions of value we did not see

plot(predictedy, t6$carval,cex=2,col="red",pch=16,
     main = 'Plot for predictedy and carval',
     xlab = 'predictedy', ylab = 'carval') #plotting them
points(predictedy,t6$carval,pch=1,cex=2,col="black")
abline(0,1)
cor(t6$carval,predictedy)^2
```
ascnlan
```{r}
#Bayes factor test
x6 <- ttestBF(t6$carval.last,t6$carval,paired = T)
x6@bayesFactor$bf
```

#Did people tend to pay more for their current car than their previous car?
```{r}
tmp1 <- cbind(mean(cardat[which(cardat$type=='Sedan'),"carval"]),mean(cardat[which(cardat$type=='Sedan'),"carval.last"]))
tmp2 <- cbind(mean(cardat[which(cardat$type=='SUV'),"carval"]),mean(cardat[which(cardat$type=='SUV'),"carval.last"]))
tmp3 <- cbind(mean(cardat[which(cardat$type=='Truck'),"carval"]),mean(cardat[which(cardat$type=='Truck'),"carval.last"]))
t7 <- rbind(tmp1,tmp2,tmp3)
row.names(t7) <- c('Sedan','SUV','Truck')
colnames(t7) <- c('carval','carval.last')
t7

ct_7 <- chisq.test(t7) 
ct_7 #Chi-squared test on t7
bf_7 <- contingencyTableBF(t7, sampleType = 'indepMulti', fixedMargin = 'cols') 
bf_7 #bayes factor contingency table tests for t7
bf_7@bayesFactor$bf  #bayes factor value
corrplot(ct_7$residuals, is.cor = F)
```

#Did trucks cost more than SUVs?
```{r}
q1<-subset( cardat, cardat$type == 'Truck') 
t8 <- cbind(sum(q1$carval),sum(q1$carval.last))

q2 <- subset( cardat, cardat$type == 'SUV') 
temp <- cbind(sum(q2$carval),sum(q2$carval.last)) #combining the sum of prices of trucks and suv
t8 <- rbind(t8,temp)
row.names(t8) <- c('Truck','SUV')
colnames(t8) <- c('carval','carval.last')
t8
```
We can see that trucks cost less than SUV in both previous price and current price.
```{r}
truck <- q1 #truck data
suv <- q2 #suv data
```

```{r}
t.test(truck$carval, suv$carval, alternative = 'less') #comparing current prices
```

```{r}
t8_bf <- ttestBF(truck$carval, suv$carval) #bayesfactor test
t8_bf
t8_bf@bayesFactor$bf
```

```{r}
t.test(truck$carval.last, suv$carval.last, alternative = 'less') #comparing previous prices
```

```{r}
t8_bf <- ttestBF(truck$carval.last, suv$carval.last) #bayesfactor test
t8_bf
t2_bf@bayesFactor$bf
```

```{r}
vioplot(truck$carval, suv$carval, names = c('truck','suv'),col = c('gold'))
title('Violinplot for current price of trucks and suv')
grid()
vioplot(truck$carval.last, suv$carval.last, names = c('truck','suv'),col = c('gold'))
title('Violinplot for previous price of trucks and suv')
grid()
```
