---
title: "PS11"
author: "Prateek Kumar prateekk@mtu.edu"
date: "December 9, 2018"
output:
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

When we speak sets of words, the more words we speak, the longer it takes.  The spokenduration data set describes data from speeded spoken duration of words. Each column decribes whether a particular word was on the list they spoke, but we will ignore those columns for the moment.  The last column tells how long it took to speak the word.

```
subject	cult	dare	fate	guess	hint	mood	oath	plea	rush	verb	zeal	time
s02	1	1	0	1	0	0	0	1	1	0	1	4.886719
s02	1	1	0	0	0	1	1	0	0	0	0	2.738281
s02	0	0	1	0	1	1	1	1	0	1	0	4.179688
s02	1	0	0	1	0	0	1	1	0	0	1	3.628906
```

Read in the data, and compute a list-length using the following function:
```{r}
data <- read.csv("spokenduration.csv") # reading the file
data$length <- as.factor(rowSums(data[,2:12])) # adding the length column
```

# 1. Categorical effect of length

First, we'd like to determine whether there is a significant relationship between list length and spoken duration. Create an anova model and test this relationship, treating list length as a factor. On the basis of the anova model, compute the effect size of this relationship. Also, compute a post-hoc Tukey HSD test and describe which differences are significant at a p=.05 level. Finally, compute the effect size for length, and describe what that means. Run both lm and aov models so you can examine the coefficients or the factor effects.
```{r}
sapply(data, class) 
```
We can see that the length column is of factor type.
```{r}
aggregate(data$time, list(data$length), mean) 
```
Calculating the means of time as per the length of words.
```{r}
library(ggpubr)

ggboxplot(data, x = "length", y = "time", 
          color = "length", palette = c("red", "gold", "green", "blue"),
          order = c("3", "4", "5", "6"),
          ylab = "Time", xlab = "Length")
```
The boxplot
```{r}
res.aov <- aov(time ~ length, data = data)
summary(res.aov)
```
As the p-value is less than the significance level 0.05, we can conclude that there are significant differences between the groups highlighted with "*" in the model summary.
```{r}
library(sjstats)
anova_stats(res.aov) #computing the effect size of the relationship
```
As the ANOVA test is significant, we can compute Tukey HSD for performing multiple pairwise-comparison between the means of groups.
```{r}
TukeyHSD(res.aov)
```
It can be seen from the output, that the difference between 5 and 3, 6 and 3, 6 and 4 is significant with an adjusted p-value of 0.0000496, 0.0, 0.0000001 respectively.
```{r}
eta_sq(res.aov, partial = T) #computing the effect size of length
```
Now calculating the effect size of length we got the value as 0.243 for partial ETA squared which is small, this means that two lengths mens does not differ by 0.243 std deviations or more, the difference is trivial, even if it is statistically signficant. 

It also means that it accounts for 24.3% of variance.
```{r}
res.lm <- lm(time ~ length, data = data)
summary(res.lm)

res.aov <- aov(time ~ length, data = data)
summary(res.aov)
```
From lm() function we can see each of the levels effect of length and aov() function gives an overall effect of length. 

We can see that length is statistically significant from the aov() result but from the result of lm() we see that all the 3 levels(3,5 and 6) are statistically significant except level 4 which is just above 0.05 value.


# 2.  Subject effects
Next, we might expect that different people speak more quickly or more slowly. We'd like to incorporate an overall speed-by-subject factor. To do so, reset the contrasts of subject to use sum-to-zero coding (so that they will all be coded with respect to the mean), and add subject to the model. Use a type-II ANOVA to test whether subject accounts for a significant proportion of variance. Compute the effect sizes (eta^2 and omega^2). Then do a post-hoc Tukey test to determine whether any individual participants were significantly faster or slower than you'd expect.  Describe your findings in words.


```{r}
contrasts(data$subject) <- contr.sum(levels(data$subject)) 
```
Resetting the contrasts of subject to use sum-to-zero coding so that they will all be coded with respect to the mean

We had seen earlier that subject is also a factor. Adding subject to the model.
```{r}
res.aov2 <- aov(time ~ length + subject, data = data)
summary(res.aov2)
```
From the ANOVA table we can conclude that both length and subject are statistically significant. These results would lead us to believe that changing length or the subject, will impact significantly the mean time.
```{r}
res.aov3 <- aov(time ~ length * subject, data = data)
summary(res.aov3)
```
It can be seen that length and subject are statistically significant, as well as their interaction.

Applying Type-II Anova
```{r}
library(car)
Anova(res.aov3, type = "II")
```
We have seen earlier that subject is statistically significant and in the Type-II Anova result we get the p-value < 0.05, hence subject is statistically significant and accounts for a significant proportion of variance.
```{r}
anova_stats(res.aov3)
anova_stats(res.aov2)
```
We can see the effect sizes (eta^2 and omega^2) from the above results.
```{r}
TukeyHSD(res.aov2)$subject
```
From the above Tukey result we got the following findings:

1. Subject 2,3,4 are faster than Subject 1
2. Subject 3,7,8 are faster than Subject 2
3. Subject 4,5,6,7,8 are faster than Subject 3
4. Subject 7,8 are faster than Subject 4
5. Subject 7,8 are faster than Subject 5
6. Subject 7,8 are faster than Subject 6 as well
7. Subject 7 is faster than Subject 8

We can derive which subjects are slower from the above statements.

# 3. Subject x length interactions
Not everyone pronounces words the same. It may be true that individuals interacts with word-length. Add the subject by length interaction to the model.  Do this in an lm and examine the coefficients.  Describe the effects--which people are slower, or which length:subject interaction terms are significantly different than the baseline? Perform a post-hoc tukey test and determine which  participants have significantly different durations for 6-item lists (e.g., the result for 6:s06-6:s01  shows whether participants s06 and s01 differed in how they pronounced 6-item lists)

Creating subject by length interaction as we did in last question
```{r}
res.lm1 <- lm(time ~ length * subject, data = data)
summary(res.lm1)
```
As the p-value is less than the significance level 0.05, we can conclude that the variables highlighted with "*" in the model summary are statistically significant.

We can see that length4,5,6 and subject1, subject3, subject7 are statistically significant and from the estimate value we can say that subject3 is slower(negative value).

With the interaction with length5 we see that only subject3 is statistically significant and with length 6 subject7,6,4 and 3 are statistically significant and we can say that those 5 interactions are significantly different than the baseline.

```{r}
TukeyHSD(aov(time ~ length * subject, data = data))$`length:subject`
```
The following participants have significantly different durations for 6-item lists:
Subject3 and 1, Subject4 and 1, Subject3 and 2, Subject7 and 2, Subject8 and 2, Subject4 and 3, Subject5 and 3, Subject6 and 3, Subject7 and 3, Subject8 and 3, Subject6 and 4, Subject7 and 4, Subject8 and 4, Subject8 and 5, Subject7 and 5.


# 4. ANCOVA
Finally, run the interaction model again, but use length as a continuous predictor instead of a categorical run both the regresion and anova models. Look at and interpret the sets of coefficients in the regression model, and interpret the results of a Type-II ANOVA. Run a post-hoc test on subject, and compute eta^2 and omega.  describe in words how you would interpret each part of this.
```{r}
class(data$length)
data$length <- as.numeric(data$length)
class(data$length)
```
Converted length to continuous predictor.
```{r}
res4.aov <- aov(time ~ length * subject, data = data) #anova model
summary(res4.aov)
```
From the anova model we can see that length and subject are statistically significant and their interaction as well.
```{r}
res4.lm <- lm(time ~ length * subject, data = data) #regression model
summary(res4.lm)
```
From the regression model we can see that length, subject1,3,4 and interaction between length and subjects 3,4,6,7 are statistically significant.
```{r}
library(car)
#Anova(res4.aov, type = "II")
Anova(res4.lm, type = "II")
```
From the Type-II ANOVA model we can see that length and subject are statistically significant and their interaction as well.
```{r warning=FALSE}
TukeyHSD(res4.aov, which = 'subject') #check for another
```
Post hoc test for 
```{r}
anova_stats(res4.aov)
```
asca
```{r}
anova_stats(res4.lm)
```
ascas