---
title: "Problem Set 1"
author: "Shane T Mueller (shanem@mtu.edu)"
output:
  html_document:
    df_print: paged
---

# Basic R Operations

For each of the following questions, provide R commands used to solve the
problem, and example output. Include both the questions and answers in
your solution. You can create an R Markdown file to turn in; you should
include the relevant code you used in the markdown, but EVERY QUESTION SHOULD BE ANSWERED IN WORDS with the code and output
just supporting that response. Unless otherwise specified, write your text as
text-based markdown, and not as code comments.


##1.  Reading and commenting code

The '#' symbol is used to write comments,
explanations and notes.  For example:
  
```
x <- 3 + 6  # Add two numbers together
```

You can use it to add sections as well:

```
##########################################
## Find the maximum of a set of ten numbers
print(max(runif(10)))
```

Read the following code, and write comments on each line explaining what is going on.


```{r}
x <- c(1,2,3,4,5,6,7) 
y <- mean(x)
x2 <- x - y  
a1 <- runif(100)
a2 <- rnorm(100)
a3 <- a1 * a2   
a.tab <- cbind(a1,a2,a3)  
pairs(a.tab)
cor(a.tab)

```



#2. Assignment of values

Determine what happens to both a and b when you do the following. Write an answer in words.

```{r}
a <- 100
b <- a
a <- 200
```
Answer:

```{r}
 a <- c(1,2,3,4,5)
 b <- a
 b[3] <- 10
```
Answer:

  
#3. Computing Means of data
The arithmetic mean of a set of numbers is the sum of those numbers divided by the number of numbers, and for a set of numbers $x$, it can be written:

```
  sum(x)/length(x)
```


There are other ways of finding the average, including the geometric mean (the Nth root of the product of N things) and the  harmonic mean (the reciprocal of the mean of the reciprocals of the numbers, where reciprocal means 1/x.). Again, for X, a way of calculating this is:

##Geometric mean (assuming all values are greater than 0):
```
exp(mean(log(x)))
exp(sum(log(x)) / length(x))
```

##Harmonic mean:

```
1 / mean(1/x)
```

For each of the  following definitions of x1..x5 below create a histogram, and then use the `abline(v=number)` function to plot the arithmetic, geometric, and harmonic means as vertical lines on the plot, and `text()` to label it.

```{r}
x1 <- runif(1000)
x2  <- runif(1000) * 2
x3 <- runif(1000) + 2
x4 <- 5/(runif(1000)+.04)
x5 <- exp(rnorm(1000))
```

That is, if you have a data set `x`, and you computed the means and saved it as `x.mean`, `x.hmean`, and `x.gmean`,  the following will make a histogram and plot the vertical lines and text labels:

```
hist(x)
abline(v=x.mean,lwd=3, col="blue")
text(x.mean,100,"Arithmetic mean")
abline(v=x.hmean,lwd=3,col="green")
text(x.hmean,100,"Harmonic mean")
abline(v=x.gmean,lwd=3,col="red")
text(x.gmean,100,"Geometric mean")

```

Try to adjust the locations of the text, etc. to make it look good.
Add at least three customizations to the hist function (i.e., change the color,
the number of breaks, the title, etc.) to make them look nice.
For each of the five data sets, describe how the different means are af-
fected by the transforms in the distribution. Suggest which means would be
appropriate for describing the middle tendency of the data


##Answers:

#4. Normalizing data.  

Sometimes you want to transform a set of numbers so they
have a mean of 0 and a standard deviation of 1.  This is done by subtracting the mean
and dividing be the standard deviation.  Normalize and plot the following befor and after,
and compute the mean and standard deviation of each normalized data set to verify
that you have normalized them.
For example, suppose you have a vector x containing unknown values:

```{r}
x <- (1:1000)[order(runif(1000))]
par(mfrow=c(1,2))
plot(x, main="before",sub=paste("mean=", mean(x),
       "sd=",round(sd(x),4)),cex=.5)
x.norm <- (x-mean(x))/sd(x)
plot(x.norm, main="after",
     sub=paste("mean=",mean(x.norm),"sd=",sd(x.norm)),cex=.5)
mean(x.norm) #[1] 0
sd(x.norm)   #[1]
```

Do the same for x1 through x5: 
```{r}

x1 <- 1:20
x2 <- runif(100) + rnorm(100)*3
x3 <- seq(1,100,3)
x4 <- c(runif(10),300)
x5 <- 1/(runif(50)*3)
```
